---
title: 'Assignment: Elliptic Curve Cryptography, Due Monday, 11:59pm'
author: "Chris Betsill"
date: "November 7, 2016"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
library(gmp)
library(cryptoCTB)
```

## 1. Is this group cyclic?

Let $E$ be the group defined by the elliptic curve $y^2 = x^3 -3x + 3$ modulo 7.

a. List all the points in $E$. Explain how you know you have found all the points.
```{r}

findPoints <- function(b, c, p){
  cat("NA, NA \n")
  for(x in 0:p-1){
    y2 <- mod.bigz(powm(x, 3, p) + mul.bigz(x, b) + c, p)
    if(y2 ==0)
      cat(x, ", 0 \n")
    if(1 == powm(y2,  div.bigz(sub.bigz(p, 1), 2), p)){
      y <- powm(y2, div.bigz(add.bigz(p, 1), 4), p)
      negy <- -y %% p
      cat(x, ",", as.integer(y), "\n")
      cat(x, ",", as.integer(negy), "\n")
    }
      
  }
}
findPoints(-3, 3, 7)
```
b. Show that the number of points in $E$ satisifies Hasse's Theorem.
```{r}
n <- 2*sqrt(7)+7+1
print(floor(n%%7))
```

From part a, you can see that there are 6 points - which according to Hasse's theorem (calculated above) is the maximum number of points on the Elliptical Curve.

c. Is the group $E$ a cyclic group? Prove or disprove.
  If all orders were prime, then it would be cyclic

## 2. How about this one?

Let $E$ be the group defined by the elliptic curve $y^2 = x^3 - x$ modulo 71.

a. Show that $E$ contains exactly 72 points.
```{r}
numPoints <- function(b, c, p){
  num <- 1
  for(x in 0:p-1){
    if(x > 0){
    y2 <- mod.bigz(powm(x, 3, p) + mul.bigz(x, b) + c, p)
    if(y2 == 0)
      num <- num + 1
    if(1 == powm(y2,  div.bigz(sub.bigz(p, 1), 2), p)){
      num <- num +2
    }
      
    }
  }
  return(num) # for some reason, the first point found has an x of -1... not sure why since x starts at 0 in the for loop. Use findPoints for details.
}
#findPoints(-1,0, 71)
numPoints(-1,0, 71)
```
b. Find all the different possible orders of elements in $E$.

```{r eval=FALSE}

#identity function runs for a while but then fails with Error in if (all(p1 == p2)) { : missing value where TRUE/FALSE needed. 
# I spent a while chasing this and was unable to find the error. If this worked, you could find orders (logic appears correct) and if all were prime, it is cyclic (part c for 1 & 2)
identity <- function(b, c, p, point){
  i <- 1
  while(1){
    pt <- ecPowModp(b, c, p, point, i)
    print(pt)
    if(all(is.na(pt)))
      return(i)
    i <- i+1
  }
}
identity(-1, 0, 71, c(2,19))
findOrders <- function(b, c, p){
  xs <- vector()
  ys <- vector()
  orders <- vector()
  for(x in 0:p-1){
    if(x >= 0){
    y2 <- mod.bigz(powm(x, 3, p) + mul.bigz(x, b) + c, p)
    if(y2 == 0){
      xs <- append(xs, x)
      ys <- append(ys, 0)
    }
    if(1 == powm(y2,  div.bigz(sub.bigz(p, 1), 2), p)){
      y <- powm(y2, div.bigz(add.bigz(p, 1), 4), p)
      negy <- -y %% p
      xs <- append(xs, x)
      ys <- append(ys, as.integer(y))
    }
      
    }
  }
  for(i in 1:length(xs))
    orders <- append(orders, identity(b, c, p, c(xs[i], ys[i])))
}
#findOrders(-1, 0, 71)
```

c. Is $E$ cyclic? Prove or disprove.

  If all orders were prime, then it would be cyclic

## 3. Find a point on a curve

Alice wants to send the message $m = 9230923203240394234$ using a cryptosystem based on the elliptic curve $y^2 = x^3 + 7x + 9$ modulo $p = 34588345934850984359911$. 

a. Show that there is no point of the form $(m, y)$ on this elliptic curve.
```{r}
m <- as.bigz("9230923203240394234")
p <- as.bigz("34588345934850984359911")
b <- 7
c <- 9

isPoint <- function(b, c, p, m){
  y2 <- mod.bigz(powm(m, 3, p)+mul.bigz(b, m)+c, p)
  if(1 == powm(y2,  div.bigz(sub.bigz(p, 1), 2), p)){
    y <- powm(y2, div.bigz(add.bigz(p, 1), 4), p)
    print("X is:")
    print(m)
    print("Y is:")
    print(y)
    return(TRUE)
  }
    
  return(FALSE)
}
isPoint(b,c,p,m)
```

b. Encode $m$ as a point on this curve by adding a digit. That is, find a point of the form $(10m+k, y)$ on this curve, for some value of $k$ between 0 and 9.

```{r}
for(i in 0:9){
  m2 <- add.bigz(i, mul.bigz(m, 10))
  if(isPoint(b,c,p,m2))
    break
}
```

c. What is the approximate probability that this method (adding a single digit) will fail to produce a point on the curve?

Approximatly $.5^{10} = 0.0009765625$

## 4. Factor using an elliptic curve

Factor the number $2875605016366351$ using Lenstra's method, with an elliptic curve of your choice. Use your `ecPowModp` function. Show your work.

E is $y^2 = x^3 + 5x -5$ mod $n$
```{r eval=FALSE}
#this code was used, and after the second (first was negative) printed statment I stopped it - I couldn't get it to break on a warning

n <- as.bigz("2875605016366351")
run <- 1
fac <- function(b, c){
for(i in 1:15){
  tryCatch(ecPowModp(b, c, n, c(1,1), factorial(i)), warning=function(w){
    cat("x^3+",b,"x+", c,"at ", i)
    print(w)
  })
}
  fac(b+1, c-1)
}
fac(1, -1)

#x^3+48x-48 at  11
```
```{r}
gcd(429877887478171, 2875605016366351) 
```

## 5. Elliptic curve discrete logarithm

Let $G = (23,14)$ be a point on the elliptic curve $y^2=x^3+4x-12063$ modulo $34543427$. Find $n$ such that $nG = (10735908, 411234)$.

```{r eval=FALSE}

# This is me trying to implement a BSGS attack on a EC discreet log. It takes a long time to run..

# The one thing I wasnt sure on from the notes is if the two values are 
g <- c(23, 14)
b <- as.bigz("4")
c <- as.bigz("-12063")
p <- as.bigz("34543427")
N <- 100000
ng <- c(as.bigz("10735908"), as.bigz("411234"))
vec1 <- vector()
vec2 <- vector()
for(j in 1:N){
  temp <- ecPowModp(b, c, p, g, j)
  append(vec1, temp)
  for(k in 1:N){
  temp2 <- ecPowModp(b, c, p, g, j)
  temp3 <- ecPowModp(b, c, p, temp2, N)
  temp4 <- ecAddModp(b, c, p, ng, ecNeg(temp3))
  vec2 <- vec2(vec2, temp4)
  if(length(intersect(vec1, vec2)) > 1){
    j <- which(vec1, intersect(vec1, vec2))
    k <-  which(vec2, intersect(vec1, vec2))
    return(j+N*k)
  }
   
  }
}
 


```

## 6. Elliptic curve ElGamal

Illustrate the ElGamal cryptosystem on the elliptic curve $y^2=x^3+4x-12063$ modulo $34543427$. Let $\alpha = G = (23,14)$. 

a. Alice wants to send the message $m=20161908$. Find a point $P_m = (m,y)$ on the curve, if possible. If no such point exists, pad $m$ to obtain a point $P_m$ on the curve.

```{r}

m <- as.bigz("20161908")
p <- as.bigz("34543427")
b <- 4
c <- -12063
pad <- 10

isPoint(b, c, p, m)

```

b. For his private key, Bob chooses $a = 1945$. What information does Bob publish?



Bob publishes the elliptical curve $E$, a point $\alpha$, and $\beta = 1945*\alpha$  

c. What message does Alice send?

Alice sends $r$ which is eqqual to $r*\alpha$ in $E$ where r is her secret random number, and $t$, which is equal to $m + k\beta$ in $E$

d. Show how Bob can decrypt the message.

Bob decrypts using his secret $a = 1945$ by computing $t - ar$ in $E$

## 7. Elliptic curve Diffie-Hellman

Illustrate the ellptic curve Diffie-Hellman key exchange on the elliptic curve $y^2=x^3+4x-12063$ modulo $34543427$. Let $G = (23,14)$. Suppose Alice's secret number is $N_A = 1984$, and Bob's secret number is $N_B = 2003$.

a. Compute the messages that Alice and Bob send to each other.
```{r}
na <- as.bigz("1984")
nb <- as.bigz("2003")
m <- as.bigz("20161908")
p <- as.bigz("34543427")
g <- c(23,14)
b <- 4
c <- -12063

ma <- ecPowModp(b, c, p, g, na)
ma
mb <- ecPowModp(b, c, p, g, nb)
mb
```

b. Compute the shared key.
```{r}
ms <- ecPowModp(b, c, p, mb, na) # note this is the same as ecPowModp(b, c, p, ma, nb)
ms
```
